---
title: "WUP_04"
author: "Daniela Olvera-Mota"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSS=read.csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  

$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The response variable is WORDSUM, it is a numerical variable.

The predictor variable is DEGREE, it is categorical variable.

The helper file that we would use for this is Numerical~Cat.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

I used ChatGPT to get R code for plotting.

```{r}
# Load required library
library(ggplot2)

# Create a box plot
ggplot(data=GSS, aes(x = DEGREE, y = WORDSUM)) +
  geom_boxplot() +
  labs(x = "Category", y = "WORDSUM") +
  ggtitle("Box Plot of Response variable by Category")
```

This graph shows the WORDSUM scores of the those with a bachelor's degree, graduate degree, high school degree, junior colleges degree, and less than high school.
  
### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=GSS)
```
This data shows the median and mean scores of WORDSUM test based on the level of education of a person has.This data shows that those who had a Junior college degree,Graduate degree, and Bachelor degree scored higher (about 9-10 score) on the WORDSUM test than those who has a high school degree or LT high school degree (about 7-8 score).Only slightly higher but there is still a significant difference.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSS)
summary.aov(model1)
```
The p value is 0.000236, which is less than 0.05.Therefore we would reject the null hypothesis which was that there was no difference between the average score on the WORDSUM test and level of schooling DEGREE.I can conclude that there is a difference in average score on the WORDSUM test based on the different levels of DEGREE a person has.


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
plot(TukeyHSD(model1))
print(TukeyHSD(model1))
```

Based on the graph we can see that the only places where there is a difference are the following :
HS to G, LTHS to G,
Everything else does not have a difference,is the same.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The original question was, is there a difference between between the scores on the WORDSUM test based on the level of schooling DEGREE a person has ?
Yes, there is a difference between the average score on the WORDSUM test and level of schooling DEGREE.

The p value was 0.000236, which is less than 0.05.Therefore we would reject the null hypothesis that there was no difference between the average score on the WORDSUM test and level of schooling DEGREE.I can conclude that there is a difference in average score on the WORDSUM test based on level of schooling degree.